---
title: "Creating the Final cleaned samples list for DE analysis"
author: "Caroline_Duncombe"
date: "2024-11-13"
output: html_document
---

PART I: DSP Analysis on GeoMX data 

Script will take DCC files and convert to normalized, batch corrected data for downstream applications. 

Step 1: Filter data based on Nanostring filter. 
Step 2: Normalize with this filtered data.
Step 3: Filter to 200um ROI (exclude the smaller ROIs).
Step 4: Batch correct with SVA method. 

########################################################################################################################################

# LOAD IN DATA

########################################################################################################################################

Load
```{r load the libraies }

library(here)
library(GeomxTools)
library(CellMembrane)
library(DESeq2)
library(Seurat)
library(dplyr)
library(ComplexHeatmap)
library(tibble)
library(sva)

#load or download the 'usethis' package.
pacman::p_load("usethis")
pacman::p_load(BiocManager, devtools)

rm(list = ls())
```

Run this if the Cell membrane token is expired:
```{r Load Cell Membrane if necessary}
if (! ("pacman" %in% installed.packages()[,"Package"])) {
  install.packages('pacman')
}
#load or download the 'usethis' package.
pacman::p_load("usethis")


usethis::create_github_token()

gitcreds::gitcreds_set()

pacman::p_load(BiocManager, devtools)

# Make sure to update your Rprofile to include Bioconductor repos, such as adding this line to ~/.Rprofile:
local({options(repos = BiocManager::repositories())})

#Latest version:
devtools::install_github(repo = 'bimberlabinternal/cellmembrane', dependencies = TRUE, upgrade = 'always')

```

Load in the DCC files 
```{r Load the datafile and create Nanostring Object}

download_directory <- here::here('raw_data/dcc_files/DCC-20240715')

  # Point the readNanoStringGeoMxSet function to the directory containing DCC files from this plate
dccFiles <- dir(download_directory, pattern = "\\w*.dcc$", full.names = TRUE)
#dccFiles <- download_directory
  
  # Point to the pkc file (kind of like a reference atlas for nanostring)
  pkcFile <- "raw_data/pkc/Mm_R_NGS_WTA_v1.0.pkc"
  
  # Point to the annotation file (metadata for the samples)
  annotationFile <- "raw_data/meta_data/annotation_data_final.xlsx"
  
  myData <- suppressWarnings(readNanoStringGeoMxSet(dccFiles = dccFiles,
                                                    pkcFiles = pkcFile,
                                                    phenoDataFile = annotationFile,
                                                    phenoDataSheet = "Master",
                                                    # this is the excel sheet’s name
                                                    phenoDataDccColName = "SAMPLE_ID",
                                                    # these are headers/column names in the excel file
                                                    protocolDataColNames = c("SegmentDisplayName", "RawReads", "unique_ID") # Add columns about sequencing data here. 
                                                    ))
  
```

########################################################################################################################################

# STEP 1: Filter out the files that failed based on the nanostring criteria. 

########################################################################################################################################


After reading in the object, we will do a couple of QC steps.

Shift all 0 counts by 1
Flag low quality ROIs
Flag low quality probes
Remove low quality ROIs and probes

```{r Filter out myData to the recommended Nanostring filters}

# Create a new object o play with
demoData <- myData


#Probe QC: Filter out probes that did not work. 
demoData <- setBioProbeQCFlags(demoData,
                    qcCutoffs=list(minProbeRatio=0.1,
                                   percentFailGrubbs=20),
                   removeLocalOutliers=FALSE)

probeflag <- featureData(demoData)[["QCFlags"]]    # no probes failed so good to go.


# ROI QC: Use the Nanostring DC flag to create a frame that has flags per ROI
demoData <- setSegmentQCFlags(demoData, 
                  qcCutoffs=list(minSegmentReads=1000, 
                                 percentAligned=80, 
                                 percentStitched=80,
                                 percentTrimmed=80,
                                 percentSaturation=50,
                                 minNegativeCount=0, 
                                 maxNTCCount=1000,
                                 minArea=1800,
                                 minNuclei=2))

demoflag <- protocolData(demoData)[["QCFlags"]]    # 5 samples failed

QCResultsIndex <- which(apply(protocolData(demoData)[["QCFlags"]], 
                              1L , function(x) sum(x) == 0L))

# Extract QC flags per ROI (5 samples failed in this case)
demoflag <- protocolData(demoData)[["QCFlags"]]

# pull out the failed values: 
failed <- rownames(demoflag)[apply(demoflag, 1, any)]

qc_dat <- protocolData(demoData)
qc_dat <- qc_dat@data

are_equal <- qc_dat$Raw == qc_dat$RawReads

QCPassed <- demoData[, QCResultsIndex]
dim(QCPassed) 

```

############# FILTERED SAMPLES #################
Final analysis includes
20175 Features (genes)
273 ROIs 

Originally 278 samples and then filtered the 5 failed samples out. 

```{r What samples failed?}

# Print the message along with the contents of the 'failed' variable
print(paste("These samples failed and should be filtered out:", failed))

#Five failed: going to filter out the dataset for only those that passed:

#[1] "These samples failed and should be filtered out: DSP-1001660018808-A-B05.dcc"
#[2] "These samples failed and should be filtered out: DSP-1001660018808-A-B08.dcc"
#[3] "These samples failed and should be filtered out: DSP-1001660018808-A-G08.dcc"
#[4] "These samples failed and should be filtered out: DSP-1001660021036-D-F12.dcc"
#[5] "These samples failed and should be filtered out: DSP-1001660021037-B-G11.dcc"

# 2/5 were connected to a 200um ROIs

```

Make the Seurat Object with the passing QC'd data:
```{r Create Seurat Object}

# Aggregate counts (for the WTA, this just aggregates negative probes)
  target_myData <- aggregateCounts(QCPassed)
  
  # The data won't export unless a normalization method is run, but I only want the raw data
  # Using for housekeeping genes. 
  target_myData <- normalize(target_myData, norm_method = "hk", fromElt = "exprs", toElt = "hk_norm")
  
  # Export the raw data to a Seurat object
  seuratObj.exp <- GeomxTools::as.Seurat(target_myData, normData = "exprs", forceRaw = TRUE)

seuratObj.play <-  seuratObj.exp
# playing with Seurat
count_dat_play <- Seurat::GetAssayData(seuratObj.exp, layer = "counts") # could add meta data to this. 

```

########################################################################################################################################

# Step 2: Normalize data with Cell Membrane

########################################################################################################################################


```{r Normalize with CellMembrane and create Seurat object}

# Normalization with CellMembrane - there is the normalization process and there is not.
# Fit a model hat is sensitive to the things that you want to take into account. 
# This does a  library sized tranformation - transcript per million reads per library.
# This is only transcripts that got aligned. 
# Seurat does per 100,000 log2 or log - see what seurat

#####

seuratObj.exp <- NormalizeData(seuratObj.exp)
seuratObj.exp <- FindVariableFeatures(seuratObj.exp, nfeatures = 5000) #5000 is completely made up
seuratObj.exp <- ScaleData(seuratObj.exp)
seuratObj.exp <- RunPCA(seuratObj.exp)

# Could review the result of the bulk data - what is the same and the spatially informed liver. 
# What is the same versus different. 

saveRDS(seuratObj.exp, file = "clean_data_manuscript/seurat_object/QC_normalized_data.rds")

```
########################################################################################################################################

# Step 3: Filter to 200um ROIs

########################################################################################################################################


Save the new dataset.
```{r Create datatable of samples that are omitted}

seuratObj.exp_200 <- subset(seuratObj.exp, ROI_size == "200um")


table(seuratObj.exp_200@meta.data$Group, seuratObj.exp_200@meta.data$Liver_zone, seuratObj.exp_200@meta.data$Type)


saveRDS(seuratObj.exp_200 , file = "clean_data_manuscript/seurat_object/QC_normalized_data_200um.rds")


```

Now down to 213 samples to include in the downstream analysis
#Now 213 samples!
There should be 9 for schizont & bystander, and 6 for mock. 

Two samples at 200um were filtered out based on a QC criteria:
*a M PC Inf on Block 1 & ORX PP Inf on block 2. 
*Never collected a M IZ on block 1 (could only find 1!) during ROI selection. 

########################################################################################################################################

# Step 4: Batch correct for the contribution of block.

########################################################################################################################################

SVA package is the final method to be applied. 

```{r SVA package}

rm(list = ls())

dat <- readRDS("clean_data_manuscript/seurat_object/QC_normalized_data_200um.rds") 

# Load in the count matrix for batch correction
gene_dat <- GetAssayData(dat, assay = "GeoMx", layer = "counts") 
bk.dat <- as.matrix(t(gene_dat))

# Create a usable dataframe for SVA by making it a matrix. 
count_matrix <- as.matrix(gene_dat)
all(colnames(count_matrix) == rownames(dat@meta.data)) # Check that columns and row match

# Making batch variable
batch <- as.factor(dat@meta.data$Block)

#Making the covariate matrix
sex <- dat@meta.data$Group
infection <- dat@meta.data$Type
covariate_matrix <-data.frame(sex = sex, infection = infection)

#check if all looks good:) the factors look good
                               
# Now run the SVA package for batch corrections with the count matrix in the pro
adjusted <- ComBat_seq(count_matrix, batch=batch, group= NULL, covar_mod = covariate_matrix)

bk.dat.adjusted <- as.matrix(adjusted)
dim(bk.dat.adjusted) # the sample-by-gene raw count matrix of bulk RNAseq expression.

# Now add a new list into Seurat hodling the SUV_adusted matrix object. 
dat[["SVA_batchCorrected"]] <- CreateAssayObject(counts = bk.dat.adjusted)

#Check that this was inputted correctly:
# Check if the new assay "SVA_batchCorrected" exists
    # Get the counts from the new assay
svacounts <- GetAssayData(dat, assay = "SVA_batchCorrected", slot = "counts") #pulling from object
print(head(svacounts)) # Checking that matrix is there. 

saveRDS(dat, file = "clean_data_manuscript/seurat_object/QC_normalized_data_200um_SVA.rds")

```

Output is the final, normalized, QC filtered, and batch corrected seurat object!

########### ALL DONE! ##############


