---
title: "5_Violin_plots_of_genes"
author: "Caroline_Duncombe"
date: "2024-11-03"
output: html_document
---

###### CAROLINE TO UPDATE FOR FINAL VERION ##########
PART V: DSP Analysis on GeoMX data 

Script will take the normalized/batch corrected data & create individual violin plot.


Working draft - no need for GW to check.

########################################################################################################################################

# STEP 1: LOAD IN DATA

########################################################################################################################################

Goal of this script is to create output violin plots for genes of interest
```{r setup, include=FALSE}

library(SEARchways)
library(BIGpicture)
library(dplyr)
library(tibble)
library(ggplot2)
library(circlize)
library(stringr)
library(RColorBrewer)
library(here)
library(readr)
library(ComplexUpset)
library(Seurat)
library(patchwork)


rm(list = ls())
here::i_am("scripts_manuscript/5_Gene_specific_Violin_plots.Rmd")
seuratObj.tmp <- readRDS(here("clean_data_manuscript/seurat_object/QC_normalized_data_200um_SVA_CCA.rds"))

```



########################################################################################################################################

# STEP 2: MAKE FUNCTIONS!!!

########################################################################################################################################

Question for GW: Do I make these violin plots with CCA corrected/integrated data or the batch corrected verison.


```{r Create the subset cat}

dat <- subset(seuratObj.tmp, subset = ROI_size == "200um" & Type != "bystander")

# Type 1
VlnPlot(object = dat, features = 'Irf7', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Irf3', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Irf9', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Ifnar1', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Ifnar2', split.by = 'Group', group.by = 'Type')

#ISGs
VlnPlot(object = dat, features = 'Ifi44', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Ifit1', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Usp18', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Ifit3', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Ifih1', split.by = 'Group', group.by = 'Type') 


#Inflammation
VlnPlot(object = dat, features = 'Cxcl9', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Cxcl10', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Saa1', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Saa2', split.by = 'Group', group.by = 'Type')
VlnPlot(object = dat, features = 'Lcn2', split.by = 'Group', group.by = 'Type') # very hormone specific. 
VlnPlot(object = dat, features = 'Msr1', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Cd74', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Ifitm3', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Ly6e', split.by = 'Group', group.by = 'Type') 

#Abortive cells:
VlnPlot(object = dat, features = 'Cxcl10', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Sqstm1', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Myc', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Hes1', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Mdm2', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Nfkbia', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Cdkn1a', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Icam1', split.by = 'Group', group.by = 'Type') 
#Could query all DEs that are more abortive and create my own Gene set enrichment analysis. 

#Other
VlnPlot(object = dat, features = 'Xcl1', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Fcer1g', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Rheb', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Fabp5', split.by = 'Group', group.by = 'Type') 


#Kuppfer cells
VlnPlot(object = dat, features = 'Clec4f', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Csf1r', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Jchain', split.by = 'Group', group.by = 'Type') 

#T cells
VlnPlot(object = dat, features = 'Cd8a', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Cxcl3', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Xcl1', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Tlr4', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Tlr7', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Gbp2', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Cd27', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Cd274', split.by = 'Group', group.by = 'Type') 
VlnPlot(object = dat, features = 'Rheb', split.by = 'Group', group.by = 'Type') 


dat <- subset(seuratObj.tmp, subset = ROI_size != "200um")

```


Create gene lists
```{r}

cell_categories <- list(
# Type 1
type1_genes = c("Irf7", "Irf3", "Irf9", "Ifnar1", "Ifnar2"),
# ISGs
isg_genes = c("Ifi44", "Ifit1", "Usp18", "Ifit3", "Ifih1"),
# Inflammation
inflammation_genes = c("Cxcl9", "Cxcl10", "Saa1", "Saa2", "Saa3", "Lcn2", "Msr1", "Cd74", "Ifitm3", "Ly6e"),
# Abortive cells
abortive_genes = c("Cxcl10", "Sqstm1", "Myc", "Hes1", "Mdm2", "Nfkbia", "Cdkn1a", "Icam1"),
# Other
other_genes = c("Xcl1", "Fcer1g", "Rheb", "Fabp5",'Camk1d',"Cd8a", "Cxcl3", "Xcl1", "Tlr4", "Tlr7", "Gbp2", "Cd27", "Cd274", "Cxcr3"),
lipid_metabolism = c("Insig1","Fabp5"),
# Kuppfer cells
kupffer_genes = c("Clec4f", "Csf1r", "Jchain", "Marco", "Mertk", "Cadm1", "Itga9", "Trpm2"), #"Masr1"
monocytes_DCs_presentation = c("H2-Aa", "H2-Eb1", "H2-Ab1", "Psap"),
leukocyte_migration_activation = c("Itk", "Txk", "Bcl11a", "Mef2c", "Bcl11b", "Satb1"),
NK_cytotoxicity = c("Cd247", "Lck", "Vav3", "Prkca"),
hepatocyte_gene = c("Alb", "Apoc3", "Apoh", "Hamp"), #"Cyp1e2"
hormones = c("Srd5a1", "Ar", "Esr1", "Esr2", "Cyp7b1", "Cyp2b13", "Hsd17b2", "Cyp7a1", "Hsd17b12")
)


# Tgtp1 - unique to females! Never called out before. 

#Cyp7b1 - related to b cell chemotasis. 

#Fasn is interesting.- fatty acid synthetase.
#Fasn is interesting.- fatty acid synthetase with Cyp4a14 interpretation. 


#Interesting point - The b cell trend switches - check the LogFC here. 


```


Dehydroepiandrosterone (DHEA) inhibits glucose 6-phosphate dehydrogenase (G6PDH) of different species and may potentially decrease intracellular glutathione. Therefore, it can have and enhance anti-parasitic action against Plasmodium spp. We evaluated the antiplasmodial activity and the interaction of DHEA with several antimalarial drugs. The inhibitory effect of DHEA on erythrocytic and G6PDH activity and changes in the content of total and reduced gluthatione Plasmodium falciparum content were also evaluated. DHEA showed antiplasmodial activity in vitro, but the potency was low (IC(50) 118.5 μM). DHEA inhibits G6PDH from healthy erythrocyte and decreases GSH content in both erythrocytes and P. falciparum. DHEA did not synergize or antagonize the antiplasmodial effect of several antimalarial drugs. The most important actions of DHEA were the inhibition of G6PDH activity, and the decrease in both P. falciparum and erythrocyte GSH. Since most of the GSH in Plasmodium spp. infected erythrocytes comes from the parasite itself, it is possible that DHEA analogs could act with higher selectivity, better potency, and might interact synergistically with antimalarials drugs.

The point is that the host response would be connected to the downstream integration of 

Dehydroepiandrosterone effect on Plasmodium falciparum and its interaction with antimalarial drugs


```{r Customize the violin plot}

# Define custom colors for each group
Group_colors <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

# Generate the violin plot with customizations
VlnPlot(object = dat, features = 'G6pc', split.by = 'Group', group.by = 'Type') +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # Set x-axis labels to 0 degrees
    legend.position = "right",                           # Position legend on the right
    axis.title.x = element_blank(),                      # Remove x-axis title
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Full box border
    axis.line = element_line(size = .5),                  # Uniform axis line thickness
    axis.ticks = element_line(size = 1)                  # Uniform tick thickness
  ) +
  scale_fill_manual(values = Group_colors)               # Apply custom colors
            # Apply custom colors

```


```{r Looping over groups}

Group_colors <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

# Set directory path for saving plots
base_save_path <- "figs/gene_violin_plots_grids"

# Function to generate and save grid plots
for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Group', group.by = 'Type') +
      theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # Set x-axis labels to 0 degrees
    legend.position = "right",                           # Position legend on the right
    axis.title.x = element_blank(),                      # Remove x-axis title
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Full box border
    axis.line = element_line(size = .5),                  # Uniform axis line thickness
    axis.ticks = element_line(size = 1)                  # Uniform tick thickness
  ) +
  scale_fill_manual(values = Group_colors)   
  })
  
  # Combine plots into a grid
  grid_plot <- wrap_plots(plots, ncol = 3)  # Adjust columns as needed
  
ggsave(filename = file.path(base_save_path, paste0(category_name, "_grid_plot.png")),
         plot = grid_plot,
         width = 12, height = 8) 
  
}


```

I want to create the panel for sex by inflammation - to emphasize different genes

```{r Figure for Sex by Inflammation Panel 5 across}


dat <- subset(seuratObj.tmp, subset = ROI_size == "200um" & Type != "bystander")

Group_colors <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

# Set directory path for saving plots
base_save_path <- "figs/gene_violin_plots_grids"

# Define cell categories with row labels (with line breaks for wrapping)
cell_categories <- list(
  "Interferon\nstimulated" = c("Irf7", "Irf3", "Irf9", "Ifit1", "Usp18"),
  "Cell stress &\nInflammation" = c("Cxcl9", "Saa3", "Lcn2", "Cd74", "Ly6e"),
  "Abortive" = c("Tgtp1", "Cxcl10", "G6pc", "Sqstm1", "Nfkbia"),
  "Metabolism" = c("Insig1", "Fabp5", "Fasn", "Cyp4a14", "Sult2a1")
)
# Function to generate and save grid plots
for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Group', group.by = 'Type') +
      theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # Set x-axis labels to 0 degrees
    legend.position = "right",                           # Position legend on the right
    axis.title.x = element_blank(),                      # Remove x-axis title
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Full box border
    axis.line = element_line(size = .5),                  # Uniform axis line thickness
    axis.ticks = element_line(size = 1)                  # Uniform tick thickness
  ) +
  scale_fill_manual(values = Group_colors)   
  })
  
  # Combine plots into a grid
  grid_plot <- wrap_plots(plots, ncol = 5)  # Adjust columns as needed
  
ggsave(filename = file.path(base_save_path, paste0(category_name, "_grid_plot_manuscript.png")),
         plot = grid_plot,
         width = 20, height = 3) 
  
}



######################## COMBINE OPTON ###############

# Loop through each category to create violin plots and arrange them
category_plots <- list()

for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene in the current category
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Group', group.by = 'Type') +
      theme(
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "right",
        axis.title.x = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        axis.line = element_line(size = 0.5),
        axis.ticks = element_line(size = 1)
      ) +
      scale_fill_manual(values = Group_colors) 
  })
  
  row_plot <- wrap_plots(plots, ncol = 5)
  
  # Create a label for the row and combine with the row plot
  row_label <- ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = category_name, fontface = "bold",angle = 0, size = 6, hjust = 0.5) +
    theme_void()
  
  # Combine the row label with the row of plots
  combined_row <- row_label + row_plot + plot_layout(widths = c(1, 10))  # Adjust widths as needed
  category_plots[[category_name]] <- combined_row
}

# Stack all labeled rows vertically and collect the legend
final_grid <- wrap_plots(category_plots, ncol = 1) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

# Save the combined plot grid
ggsave(filename = file.path(base_save_path, "gene_violin_plots_grid_manuscript.png"),
       plot = final_grid,
       width = 18, height = 10)

```


```{r Figure for Sex by Inflammation Panel}


dat <- subset(seuratObj.tmp, subset = ROI_size == "200um" & Type != "bystander")

Group_colors <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

# Set directory path for saving plots
base_save_path <- "figs/gene_violin_plots_grids"

# Define cell categories with row labels (with line breaks for wrapping)
cell_categories <- list(
  "Interferon\nstimulated" = c("Irf7", "Irf9", "Ifit1"),
  "Cell stress &\nInflammation" = c("Cxcl9", "Saa3", "Lcn2"),
  "Abortive" = c("Tgtp1", "Cxcl10", "G6pc"),
  "Metabolism" = c("Insig1", "Fabp5", "Fasn"),
  "Others" = c("Cyp4a14", "Sult2a1", "Cd74")
)


######################## COMBINE OPTON ###############

# Loop through each category to create violin plots and arrange them
category_plots <- list()

for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene in the current category
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Group', group.by = 'Type') +
      theme(
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "right",
        axis.title.x = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        axis.line = element_line(size = 0.5),
        axis.ticks = element_line(size = 1)
      ) +
      scale_fill_manual(values = Group_colors) 
  })
  
  row_plot <- wrap_plots(plots, ncol = 3)
  
  # Create a label for the row and combine with the row plot
  row_label <- ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = category_name, fontface = "bold",angle = 0, size = 6, hjust = 0.5) +
    theme_void()
  
  # Combine the row label with the row of plots
  combined_row <- row_label + row_plot + plot_layout(widths = c(1, 6))  # Adjust widths as needed
  category_plots[[category_name]] <- combined_row
}

# Stack all labeled rows vertically and collect the legend
final_grid <- wrap_plots(category_plots, ncol = 1) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

# Save the combined plot grid
ggsave(filename = file.path(base_save_path, "gene_violin_plots_grid_manuscript_3.png"),
       plot = final_grid,
       width = 12, height = 10)

```


```{r Figure of Zonation & gene}

dat <- subset(seuratObj.tmp, subset = ROI_size == "200um")
Liver_zone = c("PP" = "#006400", "PC" = "#91D1C2FF", "IZ" = "#00A087FF")



custom_colors <- list(
  Treatment = c("infected" = "orange", "uninfected" = "grey23"),  # Colors for Treatment levels
  Block = c("1" = "grey10", "2" = "grey40", "3" = "grey80"),  # Existing colors for Block
  Type = c("schizont" = "black", "bystander" = "#FFC20A", "mock" = "grey"),  # Colors for Type levels
  Group = c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF"),
  ROI_size = c("200um" = "#290916", "100um" = "#67032f", "75um" = "#a1045a","50um" = "#BE93D4"),
  Liver_zone = c("PP" = "#006400", "PC" = "#91D1C2FF", "IZ" = "#00A087FF")
)

cell_categories <- list(cyp2a4 = c("Cyp2a4"))
# Set directory path for saving plots
base_save_path <- "figs/gene_violin_plots_grids"

# Function to generate and save grid plots
for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Liver_zone', group.by = 'Group') +
      theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # Set x-axis labels to 0 degrees
    legend.position = "right",                           # Position legend on the right
    axis.title.x = element_blank(),                      # Remove x-axis title
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Full box border
    axis.line = element_line(size = .5),                  # Uniform axis line thickness
    axis.ticks = element_line(size = 1)                  # Uniform tick thickness
  ) +
  scale_fill_manual(values = Liver_zone)   
  })
  
  # Combine plots into a grid
  grid_plot <- wrap_plots(plots, ncol = 1)  # Adjust columns as needed
  
ggsave(filename = file.path(base_save_path, paste0(category_name, "_grid_plot.png")),
         plot = grid_plot,
         width = 4, height = 3) 
  
}


```



```{r Figure for Sex by Inflammation Panel M vs F only}


dat <- subset(seuratObj.tmp, subset = ROI_size == "200um" & Type != "bystander" & Group != "ORX")

Group_colors <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

# Set directory path for saving plots
base_save_path <- "figs/gene_violin_plots_grids"

# Define cell categories with row labels (with line breaks for wrapping)
cell_categories <- list(
  "Interferon\nstimulated" = c("Irf7", "Irf9", "Ifit1"),
  "Cell stress &\nInflammation" = c("Cxcl9", "Saa3", "Lcn2"),
  "Abortive" = c("Tgtp1", "Cxcl10", "G6pc")
)


######################## COMBINE OPTON ###############

# Loop through each category to create violin plots and arrange them
category_plots <- list()

for (category_name in names(cell_categories)) {
  genes <- cell_categories[[category_name]]
  
  # Create violin plots for each gene in the current category
  plots <- lapply(genes, function(gene) {
    VlnPlot(object = dat, features = gene, split.by = 'Group', group.by = 'Type') +
      theme(
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "right",
        axis.title.x = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        axis.line = element_line(size = 0.5),
        axis.ticks = element_line(size = 1)
      ) +
      scale_fill_manual(values = Group_colors) 
  })
  
  row_plot <- wrap_plots(plots, ncol = 3)
  
  # Create a label for the row and combine with the row plot
  row_label <- ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = category_name, fontface = "bold",angle = 0, size = 6, hjust = 0.5) +
    theme_void()
  
  # Combine the row label with the row of plots
  combined_row <- row_label + row_plot + plot_layout(widths = c(1, 4))  # Adjust widths as needed
  category_plots[[category_name]] <- combined_row
}

# Stack all labeled rows vertically and collect the legend
final_grid <- wrap_plots(category_plots, ncol = 1) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

# Save the combined plot grid
ggsave(filename = file.path(base_save_path, "gene_violin_plots_grid_MvF.png"),
       plot = final_grid,
       width = 10, height = 8)

```




