---
title: "20241022_Visualization of SUV and Bayes options"
author: "Caroline_Duncombe"
date: "2024-10-23"
output: html_document
---

Goal of this is to create a looping visualization of the different comparisons for each InstaPrism option outputted. 
#LOAD IN DATA
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(cluster)
library(ComplexHeatmap)
library(tibble)
library(readr)
library(InstaPrism)
library(ggpubr)
library(rstatix)
library(ggplot2)
library(reshape2)  # for melt function
library(dplyr)     # for data manipulation
library(ggpubr)    # for ggboxplot
library(ggsignif)  # for stat_pvalue_manual
library(gridExtra)  # or library(patchwork) for a different approach



#Reunning this with only the 2 plates and 2 samples. 

rm(list = ls())

```

```{r Load in phenotype data}
#Load in data for phenotypes:
dat_pheno <- readRDS("clean_data/seurat_object/normalized_data.rds") 
pheno_data <- dat_pheno@meta.data %>% rownames_to_column(var = "ddcfile")
```

```{r Load in the Insta Prism output}
# Directory where the InstaPrism results are saved
output_dir <- "clean_data/instaprism_results"

# Get the list of .rdata files in the directory
rdata_files <- list.files(output_dir, pattern = "\\.rdata$", full.names = TRUE)

# Initialize a list to store the loaded data
loaded_data <- list()

# Loop through each .rdata file and load it into the list
for (i in seq_along(rdata_files)) {
  # Load the data from the file
  load(rdata_files[i])
  
  # Store the loaded data into the list
  loaded_data[[i]] <- InstaPrism.res.initial  # Assuming InstaPrism.res.initial is the object in each file
  
  # Optionally, assign names to the loaded data based on the file names
  file_name <- basename(rdata_files[i])
  names(loaded_data)[i] <- gsub("\\.rdata$", "", file_name)  # Remove .rdata from the name
}

# Now `loaded_data` contains all the reloaded InstaPrism results

```

```{r Create a couple more options within groups of interest with 200um SUV}

#Schizont only at 200um SUV. 
names(loaded_data)

# "InstaPrism_result_dat_SUV_200"
# "InstaPrism_result_dat_SUV_200_mock"
# "InstaPrism_result_dat_SUV_200_bystander"
# "InstaPrism_result_dat_SUV_200_schizont"

```


# Part 2 - make Figures for paper

Graph 1 - Sex at baseline (mock 200um)
immune_cells
structure_cells

Graph 2 - zonation (200um all)
hepatocytes_zone

Graph 3 - Bystander & Peripheral  (200um all)
immune_cells
structure_cells
hepatocytes_activation

Graph 4 - Sex on schizont  (200um schizont)
immune_cells
structure_cells
hepatocytes_activation

```{r Cell categories}

hepatocytes_zone <- c("Pericentral hepatocytes","Periportal hepatocytes", "Midlobular hepatocytes")

hepatocytes_activation <- c("Activated periportal hepatocytes","Inflammatory hepatocytes")

immune_cells <- c("Monocytes & DCs","Kupffer cells","T & NK cells", "B cells")

structure_cells <- c("Cholangiocytes","Endothelial cells","Fibroblasts", "Mesothelial cells")

other_cells <- c("Periportal endothelial cells" )

```

```{r Set up boxplot colors}

Group_colors  <- c("F" = "#941751", "M" = "#005493", "ORX" = "#52B2BF")

Zone_colors = c("PP" = "#006400", "PC" = "#91D1C2FF", "IZ" = "#00A087FF")

Type_colors = c("schizont" = "black", "bystander" = "#FFC20A", "mock" = "grey")
```

```{r Step 1 - Function to create boxplots of the data for each unique loaded data and save into a file}


plot_gene_boxplot_custom <- function(cell_type_rowname, x_axis_column, Instaprism_data, pheno_data, manual_colors) {  
  # Extract the cell data from the S4 object
  

  # Check if the specified cell type is present in the cell data
  if (!(cell_type_rowname %in% rownames(cell_data))) {
    stop("The specified cell type is not present in the Instaprism_data.")
  }

  # Extract the row of interest (cell type)
  cell_row <- cell_data[cell_type_rowname, , drop = FALSE]

  # Reshape cell data to long format (dccfile, value)
  cell_long <- melt(cell_row, variable.name = "ddcfile", value.name = "expression")

  # Filter phenotype data to only include matching samples (dccfiles)
  names_in_comparison <- unique(cell_long$Var2)
  pheno_data_to_apply <- pheno_data %>% filter(ddcfile %in% names_in_comparison)

  # Join proportion matrix data with phenotype data based on dccfile
  plot_data <- pheno_data_to_apply %>%
    left_join(cell_long, by = c("ddcfile" = "Var2")) %>%  # Ensure correct joining
    mutate(labels = .data[[x_axis_column]])

  # Check if the number of unique groups is either 2 or 3
  unique_groups <- length(unique(plot_data$labels))
  if (unique_groups == 2 || unique_groups == 3) {

    # Perform statistical test based on the number of groups
    if (unique_groups == 3) {
      stat.test <- plot_data %>%
        dunn_test(expression ~ labels, p.adjust.method = "bonferroni") %>%
        add_significance()
      plot_width <- 4.5
      plot_height <- 4.5
    } else if (unique_groups == 2) {
      stat.test <- plot_data %>%
        wilcox_test(expression ~ labels) %>%
        add_significance()
      plot_width <- 2.5
      plot_height <- 3
    }

    # Annotate p-values with significance labels
    stat.test <- stat.test %>%
      mutate(p_label = case_when(
        p.adj > 0.1 ~ "ns",
        p.adj <= 0.1 & p.adj > 0.05 ~ as.character(round(p.adj, 3)),
        p.adj <= 0.05 & p.adj > 0.01 ~ "*",
        p.adj <= 0.01 & p.adj > 0.001 ~ "**",
        p.adj <= 0.001 ~ "***"
      ))

    # Calculate dynamic y positions for significance annotations
    max_y_value <- max(plot_data$expression, na.rm = TRUE)
    y_positions <- seq(max_y_value * 1.0, max_y_value * 1.2, length.out = nrow(stat.test))
# Define a buffer for the y-axis
    # Plot the data
    p <- ggboxplot(plot_data, x = "labels", y = "expression",  
               add = "mean_se", fill = "labels") +
  geom_jitter(position = position_jitter(width = 0.2, height = 0), show.legend = FALSE) +
  stat_pvalue_manual(stat.test, label = "p_label", hide.ns = FALSE, 
                     y.position = y_positions, size = 5, label.size = 5, steps = 1, tip.length = 0) +
  labs(
    title = paste(cell_type_rowname),
    x = NULL,
    y = "Expression Level"
  ) +
  scale_fill_manual(values = manual_colors) +  # Apply custom colors
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_classic() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    panel.border = element_rect(color = "black", fill = NA, size = 1),  # Add box border
    axis.ticks.y.right = element_line(size = 1),  # Right axis ticks
    axis.ticks.x.top = element_line(size = 1),    # Top axis ticks
    axis.line.y.right = element_line(size = 1),    # Right axis line
    axis.line.x.top = element_line(size = 1)       # Top axis line
  )

  } else {
    stop("The function is designed for datasets with either two or three unique groups.")
  }

  return(p)  # Return the plot object
}

# Example call
plot_gene_boxplot_custom("Kupffer cells", "Group", Instaprism_data, pheno_data, Group_colors)

```

```{r PLot for combined boxplots}
plot_combined_gene_boxplots <- function(cells_of_interest, x_axis_column, grid_columns, Instaprism_data, pheno_data, colors, save_name, width_use, height_use) {
  # Create a folder to save combined plots if it doesn't exist
  combined_folder <- "figs/Bayes_Prism_boxplots/paper_boxplots/"
  if (!dir.exists(combined_folder)) {
    dir.create(combined_folder)
  }
  
  # Prepare to store individual plots
  plot_list <- list()
  
  # Loop through each cell of interest to create plots
  for (cell_type in cells_of_interest) {
    p <- plot_gene_boxplot_custom(cell_type, x_axis_column, Instaprism_data, pheno_data, colors)
    plot_list[[cell_type]] <- p  # Store the plot in the list
  }
  
    combined_plot <- wrap_plots(plot_list, ncol = grid_columns) + 
    plot_layout(guides = "collect") +  # Collect legends
    theme(legend.position = "topright")  # Position the legend in the upper right
  
  # Add one legend for the combined plot
  combined_plot <- combined_plot + 
    guides(fill = guide_legend(nrow = 1))

  # Save the combined plot
  ggsave(filename = file.path(combined_folder, save_name), 
         plot = combined_plot, 
         width = width_use, height = height_use)  # Adjust width and height as needed
}

```


```{r Figure on Mock, Sex}

# Define the source group for the graphs: 
cells_of_interest <- c(immune_cells, structure_cells)
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_mock # Just mock samples
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
pheno_data <- pheno_data
colors <- Group_colors


# Call the function to generate and save combined boxplots
plot_combined_gene_boxplots(cells_of_interest, "Group", 4, Instaprism_data, pheno_data, colors, "Sex_baseline_combined_boxplots.png", 12, 6)

```

```{r Figure on Zone}

# Define the source group for the graphs: 
cells_of_interest <- c(hepatocytes_zone)
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200 # all samples
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
pheno_data <- pheno_data
colors <- Zone_colors


# Call the function to generate and save combined boxplots
plot_combined_gene_boxplots(cells_of_interest, "Liver_zone", 3, Instaprism_data, pheno_data, colors, "Zone_all_combined_boxplots.png", 10, 3)

```

```{r Figure on Type (Infection status)}

# Define the source group for the graphs: 
cells_of_interest <- c(immune_cells, structure_cells,hepatocytes_activation)
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200 # all samples
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
pheno_data <- pheno_data
colors <- Type_colors


# Call the function to generate and save combined boxplots
plot_combined_gene_boxplots(cells_of_interest, "Type", 5, Instaprism_data, pheno_data, colors, "Type_all_combined_boxplots.png", 18, 7)

```

```{r Figure on Sex in Schizont (Infection status)}

# Define the source group for the graphs: 
cells_of_interest <- c(immune_cells, structure_cells, hepatocytes_activation)
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_schizont # all samples
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
pheno_data <- pheno_data
colors <- Group_colors


# Call the function to generate and save combined boxplots
plot_combined_gene_boxplots(cells_of_interest, "Group", 5, Instaprism_data, pheno_data, colors, "group_on schizont_combined_boxplots.png", 18, 7)

```

```{r Figure on Sex in Bystander (Infection status)}

# Define the source group for the graphs: 
cells_of_interest <- c(immune_cells, structure_cells, hepatocytes_activation)
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_bystander # all samples
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
pheno_data <- pheno_data
colors <- Group_colors


# Call the function to generate and save combined boxplots
plot_combined_gene_boxplots(cells_of_interest, "Group", 5, Instaprism_data, pheno_data, colors, "group_on Bystander_combined_boxplots.png", 18, 7)

```




#Part 2 - create boxplots for all comparisons to look at trends

```{r Step 3 - Function to create final boxplots}


plot_gene_boxplot <- function(cell_type_rowname, x_axis_column, Instaprism_data, pheno_data) {
  
  # Extract the cell data from the S4 object
  cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)
  
  # Check if the specified cell type is present in the cell data
  if (!(cell_type_rowname %in% rownames(cell_data))) {
    stop("The specified cell type is not present in the Instaprism_data.")
  }
  
  # Extract the row of interest (cell type)
  cell_row <- cell_data[cell_type_rowname, , drop = FALSE]
  
  # Reshape cell data to long format (dccfile, value)
  cell_long <- melt(cell_row, variable.name = "ddcfile", value.name = "expression")
  
  # Filter phenotype data to only include matching samples (dccfiles)
  names_in_comparison <- unique(cell_long$Var2)
  pheno_data_to_apply <- pheno_data %>% filter(ddcfile %in% names_in_comparison)
  
  # Join proportion matrix data with phenotype data based on dccfile
  plot_data <- pheno_data_to_apply %>%
    left_join(cell_long, by = c("ddcfile" = "Var2")) %>%  # Ensure correct joining
    mutate(labels = .data[[x_axis_column]])
  
  # Check if the number of unique groups is either 2 or 3
  unique_groups <- length(unique(plot_data$labels))
  if (unique_groups == 2 || unique_groups == 3) {
    
    # Perform statistical test based on the number of groups
    if (unique_groups == 3) {
      stat.test <- plot_data %>%
        dunn_test(expression ~ labels, p.adjust.method = "bonferroni") %>%
        add_significance()
      plot_width <- 4.5
      plot_height <- 4.5
    } else if (unique_groups == 2) {
      stat.test <- plot_data %>%
        wilcox_test(expression ~ labels) %>%
        add_significance()
      plot_width <- 2.5
      plot_height <- 3
    }
    
    # Annotate p-values with significance labels
    stat.test <- stat.test %>%
      mutate(p_label = case_when(
        p.adj > 0.1 ~ "ns",
        p.adj <= 0.1 & p.adj > 0.05 ~ as.character(round(p.adj, 3)),
        p.adj <= 0.05 & p.adj > 0.01 ~ "*",
        p.adj <= 0.01 & p.adj > 0.001 ~ "**",
        p.adj <= 0.001 ~ "***"
      ))
    
    # Calculate dynamic y positions for significance annotations
    max_y_value <- max(plot_data$expression, na.rm = TRUE)
    y_positions <- seq(max_y_value * 1.05, max_y_value * 1.2, length.out = nrow(stat.test))
    
    # Plot the data
    p <- ggboxplot(plot_data, x = "labels", y = "expression", 
                   add = "mean_se", fill = "labels") +
      geom_jitter(position = position_jitter(width = 0.2, height = 0), show.legend = FALSE) +
      stat_pvalue_manual(stat.test, label = "p_label", hide.ns = FALSE, 
                         y.position = y_positions, size = 5, label.size = 5) +
      labs(
        title = paste(cell_type_rowname),
        x = NULL,
        y = "Expression Level"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)
      )
    
  } else {
    stop("The function is designed for datasets with either two or three unique groups.")
  }
}

```

```{r Step 2 - Loop over phenotype of interest to folder}

# List of phenotypes of interest
phenotype_of_interest <- c("Group", "Type", "Block", "Liver_zone")

# Create the output directory if it doesn't exist
output_dir <- "figs/Bayes_Prism_boxplots"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

# Loop through each dataset in loaded_data
for (data_name in names(loaded_data)) {
  
  # Access the specific InstaPrism results object
  InstaPrism_res <- loaded_data[[data_name]]
  
  # Extract the cell data matrix from InstaPrism results
  cell_data <- as.matrix(InstaPrism_res@Post.ini.cs@theta)
  
  # Loop through each phenotype of interest
  for (phenotype in phenotype_of_interest) {
    
    # Initialize a list to store plots for the current phenotype
    plots <- list()
    
    # Loop through each row (gene/cell type) in the cell_data
    for (cell_name in rownames(cell_data)) {
      
      # Try generating and saving the boxplot, handle errors gracefully
      try({
        # Generate the boxplot for each gene using the updated function
        p <- plot_gene_boxplot(
          cell_type_rowname = cell_name,  # Corrected variable name
          x_axis_column = phenotype,        # Use the phenotype as the x-axis grouping
          Instaprism_data = InstaPrism_res, # Pass the correct Instaprism data
          pheno_data = pheno_data           # Phenotype data
        )
        
        # Add the plot to the list with a unique name
        plots[[paste(data_name, cell_name, phenotype, sep = "_")]] <- p
        
      }, silent = TRUE)  # Use silent = TRUE to skip over any errors without breaking the loop
    }  # End cell_name loop
    
    # Check if there are plots to combine
    if (length(plots) > 0) {
      # Combine all plots into a grid
      combined_plot <- grid.arrange(grobs = plots, ncol = 7)  # Adjust ncol for layout

      # Create a subdirectory for the current phenotype if it doesn't exist
      phenotype_folder <- file.path(output_dir, phenotype)
      if (!dir.exists(phenotype_folder)) {
        dir.create(phenotype_folder, recursive = TRUE)
      }
      
      # Save the combined plot as a PNG file in the phenotype folder
      combined_filename <- paste0(phenotype_folder, "/", data_name, "_combined_boxplots.png")
      ggsave(combined_filename, plot = combined_plot, width = 30, height = 8)  # Adjust dimensions as needed
    }
  }  # End phenotype loop
}  # End dataset loop

```

Conclusion: The data looks good with the SUV correction. The samples I would expect to be up are up and the trends are there. Block 3 seems to be a little off, but combined the trends are there! I can work with this. 

# Part 3 - do math to include in paper on cell proportions

```{r Quantifying the proportion of hepatocytes}

# Select dataset
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_mock

# Define lists for hepatocytes and non-hepatocytes
hepatocyte <- c(hepatocytes_zone, hepatocytes_activation)
non_hepatocyte <- c(immune_cells,structure_cells,other_cells)

# Convert cell_data to a matrix if it’s not already one (as you've done)
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)

# Sum rows for hepatocytes and non-hepatocytes
hepatocyte_sum <- rowMeans(cell_data[hepatocyte, , drop = FALSE])
hepatocyte_range <- apply(cell_data[hepatocyte, , drop = FALSE], 1, range)

non_hepatocyte_sum <- rowMeans(cell_data[non_hepatocyte, , drop = FALSE])
non_hepatocyte_range <- apply(cell_data[non_hepatocyte, , drop = FALSE], 1, range)


# Calculate sums for upper and lower limits
hepatocyte_total <- sum(hepatocyte_sum)  # Upper limits
non_hepatocyte_total <- sum(non_hepatocyte_sum) 

# Calculate sums for upper and lower limits
hepatocyte_range_sum_upper <- sum(hepatocyte_range[2, ])  # Upper limits
hepatocyte_range_sum_lower <- sum(hepatocyte_range[1, ])  # Lower limits

non_hepatocyte_range_sum_upper <- sum(non_hepatocyte_range[2, ])  # Upper limits for non-hepatocytes
non_hepatocyte_range_sum_lower <- sum(non_hepatocyte_range[1, ])  # Lower limits for non-hepatocytes

# Print results
cat("Hepatocyte Proportion - Average:", hepatocyte_total, "\nRange:", hepatocyte_range_sum_lower, hepatocyte_range_sum_upper, "\n")
cat("Non-Hepatocyte Proportion - Average:", non_hepatocyte_total, "\nRange:", non_hepatocyte_range_sum_lower, non_hepatocyte_range_sum_upper, "\n")


```


```{r Now create a graph for the supplements}

Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_mock

# Create a data frame for plotting
data_plot <- data.frame(
  Cell_Type = c("Hepatocytes", "Non-Hepatocytes"),
  Average = c(hepatocyte_total, non_hepatocyte_total),
  Lower_Range = c(hepatocyte_range_sum_lower, non_hepatocyte_range_sum_lower),
  Upper_Range = c(hepatocyte_range_sum_upper, non_hepatocyte_range_sum_upper)
)

# Reshape the data for plotting
data_long <- tidyr::pivot_longer(data_plot, cols = c("Lower_Range", "Upper_Range"),
                                  names_to = "Range_Type", values_to = "Value")

# Create the bar plot
ggplot(data_plot, aes(x = Cell_Type, y = Average, fill = Cell_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Lower_Range, ymax = Upper_Range), width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Average Proportions of Hepatocytes vs. Non-Hepatocytes",
       x = "Cell Type",
       y = "Average Proportion") +
  theme_minimal() +
  theme(legend.position = "none")

```


```{r attempt}

hepatocyte <- c(hepatocytes_zone, hepatocytes_activation)
non_hepatocyte <- c(immune_cells,structure_cells,other_cells)

Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_mock
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)

# Reshape cell data to long format (dccfile, value)
cell_long <- melt(cell_data, variable.name = "ddcfile", value.name = "expression")

# Filter phenotype data to only include matching samples (dccfiles)
names_in_comparison <- unique(cell_long$Var2)
pheno_data_to_apply <- pheno_data %>% filter(ddcfile %in% names_in_comparison)

plot_data <- pheno_data_to_apply %>%
    left_join(cell_long, by = c("ddcfile" = "Var2")) %>%
  mutate(hepatocyte_des = case_when (Var1 = hepatocyte ~ "hepatocyte",
                                 Var1 = non_hepatocyte ~ non_hepatocyte ~ "Not hepatocyte"))


ggplot(plot_data, aes(x = hepatocyte_des, y = expression, fill = hepatocyte_des)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Lower_Range, ymax = Upper_Range), width = 0.2, position = position_dodge(0.9)) +
  facet_wrap(~Group)
  labs(title = "Average Proportions of Hepatocytes vs. Non-Hepatocytes",
       x = "Cell Type",
       y = "Average Proportion") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine hepatocyte and non-hepatocyte identifiers
hepatocyte <- c(hepatocytes_zone, hepatocytes_activation)
non_hepatocyte <- c(immune_cells, structure_cells, other_cells)

# Load the InstaPrism data
Instaprism_data <- loaded_data$InstaPrism_result_dat_SUV_200_mock
cell_data <- as.matrix(Instaprism_data@Post.ini.cs@theta)

cell_long <- melt(cell_data, variable.name = "ddcfile", value.name = "expression")

# Filter phenotype data to only include matching samples (dccfiles)
names_in_comparison <- unique(cell_long$Var2)
pheno_data_to_apply <- pheno_data %>% filter(ddcfile %in% names_in_comparison)

# Merge phenotype and cell data
plot_data <- pheno_data_to_apply %>%
  left_join(cell_long, by = c("ddcfile" = "Var2")) %>%
  mutate(hepatocyte_des = case_when(
    Var1 %in% hepatocyte ~ "Hepatocyte",
    Var1 %in% non_hepatocyte ~ "Non-Hepatocyte",
    TRUE ~ "Other"
  ))


# Create the plot with jitter and error bars
ggbarplot(data = plot_data, 
          x = "Var1", 
          y = "expression", 
          fill = "hepatocyte_des", 
          add = "mean_se",  # Add mean and standard error bars
          error.plot = "errorbar",
          width = 0.7) +  # Adjust bar width
  geom_jitter(aes(color = Block)) +  # Add jitter for individual points
  facet_wrap(~ Treatment) +
  labs(title = "Average Proportions of Hepatocytes vs. Non-Hepatocytes",
       x = "Cell Type",
       y = "Average Proportion") +
  theme_minimal() +
  theme(legend.position = "right")

# Just hepaatocytes

summary_data <- plot_data %>%
  group_by(hepatocyte_des, ddcfile, Group, Block) %>%  # Include Block in grouping
  summarise(sum_expression = sum(expression, na.rm = TRUE))

ggbarplot(data = summary_data, 
          x = "hepatocyte_des", 
          y = "sum_expression", 
          fill = "hepatocyte_des", 
          add = "mean_se",  # Add mean and standard error bars
          error.plot = "errorbar",
          width = 0.7) +  # Adjust bar width
  geom_jitter(aes(color = Block)) +  # Add jitter for individual points
  facet_wrap(~ Group) +
  labs(title = "Average Proportions of Hepatocytes vs. Non-Hepatocytes",
       x = "Cell Type",
       y = "Average Proportion") +
  theme_minimal() +
  theme(legend.position = "right")



```
